---
title: "MOFA analysis of multi-omics CLL data"
output: 
  BiocStyle::html_document:
    code_folding: hide
    toc: yes
    toc_depth: 5
    toc_float: yes
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: '3'
editor_options:
  chunk_output_type: console
---

# Setup
## load packages
```{r load packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(MultiAssayExperiment)
library(MOFA)
library(pheatmap)
library(UpSetR)
library(reticulate)
library(DESeq2)
library(ggpubr)
```

## paths & environments
```{r specify python environment}
# specify conda environment to ensure python and required modules are available
use_condaenv(condaenv = "mofaEnvironment", required = TRUE)
```

```{r output path}
model_output_path <- file.path(getwd(),"data/MOFA_models")
```


## load data
```{r load data}
load(file.path(getwd(),"data/multiomics_MAE_CLL.RData"))
multiomics_MAE <- multiomics_MAE_CLL
```

```{r load gene sets}
# all gene sets were downloaded from MSigDB (v7.0)
# Subramanian, Tamayo, et al. (2005, PNAS 102, 15545-15550)

# load gene ontology gene sets from MSig db with hgnc identifiers
load(file.path(getwd(),"data/GO_gene_sets.RData"))
```


## general plot themes
```{r general plot themes}
source(file.path(getwd(),"Figure_layouts.R"))
```

## function for IGHV/tri12 comparison plot
```{r factor scatter plotting function}
plotFactorComparison <- function(factors = NULL,
                                 model_factors = NULL){
  IGHV_tri12_df <- model_factors %>% as.data.frame() %>% 
                    dplyr::select(factors) %>% 
                    rownames_to_column(var = "patient_ID") %>% 
                    left_join(.,
                            dplyr::select(as.data.frame(colData(
                              multiomics_MAE)), 
                              c("patient_ID", "trisomy12",
                                "IGHV")), by = "patient_ID") %>%
                    column_to_rownames(var="patient_ID")

  IGHV_tri12_df <- IGHV_tri12_df %>% mutate(status = 
                    ifelse(trisomy12==1 & IGHV == "M", "trisomy12&IGHVmutated",
                    ifelse(trisomy12 == 0 & IGHV == "M","IGHVmutated",
                    ifelse(trisomy12 == 1 & IGHV == "U","trisomy12",
                    ifelse(trisomy12 == 0 & IGHV == "U","IGHVunmutated",
                           NA))))) %>%
                    filter(!is.na(status))
                    


  ggscatter(IGHV_tri12_df, x = factors[1], y = factors[2], fill = "status",
            color = "status", palette = c("#bebada", "#fb8072","#8dd3c7",
                                          "#ffff4e"), 
            shape = 21, size = 3.5,
            ellipse = FALSE) +
    theme(legend.position = "right")
  }
```


# Inspect data object
```{r inspect data object}
MultiAssayExperiment::upsetSamples(multiomics_MAE)

# # display parts ot the data object
# experiments(multiomics_MAE)
# sampleMap(multiomics_MAE)
# metadata(multiomics_MAE)

knitr::kable(colnames(colData(multiomics_MAE)), format = "markdown", 
             caption = "Patient data", col.names = "patient_metadata")
```

# Preprocess data
## RNA seq & proteomcis IDs
```{r joint IDs}
# find joint features between proteomics and RNA seq
proteomics_IDs <- as.data.frame(multiomics_MAE@ExperimentList$proteomics)%>% 
                    rownames_to_column(.) %>% 
                    left_join(., 
                              multiomics_MAE@metadata$gene_symbol_mapping, 
                              by = c("rowname" = "hgnc_symbol")) %>%
                    .[,c("rowname","ensembl_gene_id")]

RNAseq_IDs <- as.data.frame(multiomics_MAE@ExperimentList$RNAseq_norm) %>% 
                rownames_to_column(.) %>%
                left_join(., multiomics_MAE@metadata$gene_symbol_mapping,
                          by = c("rowname" = "ensembl_gene_id")) %>%
                .[,c("rowname","hgnc_symbol")]

joint_IDs <- inner_join(proteomics_IDs, RNAseq_IDs, 
                    by = c("ensembl_gene_id" = "rowname"))


# filter for the joint IDs and make hgnc symbol the general identifier
multiomics_MAE@ExperimentList$proteomics <- 
                              multiomics_MAE@ExperimentList$proteomics %>%
                              .[c(joint_IDs$hgnc_symbol),]

multiomics_MAE@ExperimentList$RNAseq_norm <- 
                              multiomics_MAE@ExperimentList$RNAseq_norm %>%
                              .[c(joint_IDs$ensembl_gene_id),] %>%
                              as.data.frame() %>%
                              rownames_to_column() %>% 
                              left_join(.,joint_IDs,
                                        by = c("rowname" = 
                                                 "ensembl_gene_id")) %>%
                              column_to_rownames(., var = "hgnc_symbol") %>%
                              select(-c("rowname", "rowname.y"))
```


## Proteomics
Proteins which were not detected in 10 or more patients were excluded. Only the top 2000 proteins with the highest variance across samples were considered for further analyses.  
```{r proteomics preprocessing}
proteomics_df <- as.matrix(multiomics_MAE@ExperimentList$proteomics)

#drop all features with more than 10 NAs
proteomics_df <- proteomics_df[rowSums(is.na(proteomics_df)) <= 10,]

# keep top 2000 proteins with highest variance
proteomics_vars <- rowVars(proteomics_df, na.rm = TRUE)
proteomics_df <- proteomics_df[order(proteomics_vars,
                                     decreasing = T)[1:2000],]

hist(rowVars(proteomics_df, na.rm = TRUE), breaks  = 2000,
     main = "Variance of proteomics data after filtering", ylim = c(0,100),
     xlab = "variance of protein abundance")

multiomics_MAE@ExperimentList$proteomics <- proteomics_df
```


## RNA seq
Transcripts with a mean normalized abundance of 200 or lower across all patients were excluded. Size factors were already calculated. Variance stabilization transformation was performed. The 2000 transcripts with the highest variance are considered for further analyses. Replicated samples were averaged.
```{r rna seq preprocessing}
RNAseq_df <- as.matrix(assays(multiomics_MAE)$RNAseq_norm)

# keep top 2000 rows with the highest variance 
RNAseq_vars <- rowVars(RNAseq_df)
RNAseq_df <- RNAseq_df[order(RNAseq_vars, decreasing = T)[1:2000],]
pheatmap(RNAseq_df, cluster_cols = FALSE, cluster_rows = TRUE, 
         show_rownames = FALSE, na_col = "dark grey")

multiomics_MAE@ExperimentList$RNAseq_norm <- RNAseq_df
```


## Drug screen data
Viability values over 1.5 were removed and considered as outliers. rowMeans and rowVars cut offs were applied. If a drug at a certain concentration showed a mean viability under 0.3, it was removed from the data set. In addition, drug-concentration-combination were excluded from this analysis, if the mean viability was high (0.9-1.5), but the variance was low (equal or smaller than 0.02). 
```{r preprocess drug screen data, message=FALSE}
# change β to beta in rownames &
# set threshold of 1.5 and replace values above with NA
# apply rowMeans and rowVars cut offs

drug_resp <- assays(multiomics_MAE)$drug_resp_mono
rownames(drug_resp) <- gsub("\u03b2", "beta", rownames(drug_resp))
drug_resp[drug_resp >= 1.5] <- NA
drug_resp$rowvars <-  rowVars(as.matrix(drug_resp), na.rm = TRUE)
drug_resp$rowmeans <- rowMeans(as.matrix(drug_resp), na.rm = TRUE)

drug_resp <- drug_resp %>% rownames_to_column() %>% 
                   filter(rowmeans > 0.3) 
drug_resp <- drug_resp %>% 
                    filter(between(rowmeans, 0.9, 1.5)) %>%
                    filter(rowvars <= 0.02) %>% 
                    anti_join(drug_resp,.,)
drug_resp <- drug_resp %>% select(-c("rowvars", "rowmeans")) %>% 
                    column_to_rownames()
multiomics_MAE@ExperimentList$drug_resp_mono <- drug_resp
```

## create genetics set
This data set combines SNPs, fish data, complex aberrant caryotype and IGHV mutation status. 
```{r create genetics data sets, warning=FALSE}
genetics_df <- list(multiomics_MAE@ExperimentList$SNPs,
                    multiomics_MAE@ExperimentList$chrom_abber,
                    multiomics_MAE@ExperimentList$health_record_bin[
                      c("komplex_abberant_karyotype", "IGHV_mutated"),])%>%
  lapply(., function(df){rownames_to_column(as.data.frame(df))}) %>% 
  bind_rows(.) %>% 
  column_to_rownames(., var = "rowname")
genetics_df <- genetics_df[, c(colnames(
                multiomics_MAE@ExperimentList$health_record_bin))]

multiomics_MAE <- c(multiomics_MAE, genetics = genetics_df, mapFrom = 5L)
```

## create clinical feature set
This data set includes treatment status, gender and age category. 
```{r create set with clinical features, warning=FALSE}
clinical_features_df <- multiomics_MAE@ExperimentList$health_record_bin[c(
                          "treated", "gender_binary", "elderly_patient"),]
clinical_features_df <- clinical_features_df[, c(colnames(
                          multiomics_MAE@ExperimentList$health_record_bin))]

multiomics_MAE <- c(multiomics_MAE, 
                    clinical_features = clinical_features_df,
                    mapFrom = 5L)
```


# MOFA analysis
## Prepare MOFA analysis
```{r }
proteomics_noDrug_obj <- multiomics_MAE[,,c("genetics", "proteomics", 
                                            "RNAseq_norm")]
proteomics_noDrug_obj <- createMOFAobject(proteomics_noDrug_obj) 
plotDataOverview(proteomics_noDrug_obj)
```

```{r }
TrainOptions <- getDefaultTrainOptions()
ModelOptions <- getDefaultModelOptions(proteomics_noDrug_obj)
DataOptions <- getDefaultDataOptions()

TrainOptions$DropFactorThreshold <- 0.015 # % variance explained by factor
TrainOptions$tolerance <- 0.01 # 0.01 is recommended
TrainOptions$seed <- 2019 #random number
```

## Model creation/selection
```{r }
# ## model loop: run once -> afterwards just load models (see below)
# 
# n_inits <- 10 # number of interations
# model_list_proteomics_noDrug <- lapply(seq_len(n_inits), function(it) {
#                            TrainOptions$seed <- 2019 + it
# 
#                             MOFAobject <- prepareMOFA(
#                               proteomics_noDrug_obj,
#                               DataOptions = DataOptions,
#                               ModelOptions = ModelOptions,
#                               TrainOptions = TrainOptions)
#                             runMOFA(MOFAobject,
#                                     outfile = paste0(model_output_path,
#                                                      "/", Sys.Date(),
#                                             "_proteomics_model_noDrug_",
#                                                      it,
#                                                      ".hdf5"))
#                             })
# 
# save(model_list_proteomics_noDrug,
#      file = paste0(model_output_path, "/proteomics_models_noDrug_list.RData"))
# 

# loads the saved model list
load(paste0(model_output_path, "/proteomics_models_noDrug_list.RData"))

compareModels(model_list_proteomics_noDrug) + pp_sra_noguides_tilted
compareFactors(model_list_proteomics_noDrug)

# select the best model
proteomics_model_noDrug <- selectModel(model_list_proteomics_noDrug, 
                                       plotit = FALSE)
print(paste0("model ", proteomics_model_noDrug@TrainOptions$seed - 2019, 
             " selected"))

# save factors of chosen model (loadings from patients)
proteomics_model_noDrug_factors <- getFactors(proteomics_model_noDrug,
                                       factors = "all",
                                       as.data.frame = FALSE)
save(proteomics_model_noDrug_factors,
     file = paste0(model_output_path, 
                   "/proteomics_model_noDrug_factors.RData"))

# save weights of chosen model (loadings from features)
proteomics_model_noDrug_weights <- getWeights(proteomics_model_noDrug,
                                       factors = "all",
                                       as.data.frame = FALSE)
save(proteomics_model_noDrug_weights,
     file = paste0(model_output_path, 
                   "/proteomics_model_noDrug_weights.RData"))
```

## Inspect model
```{r , fig.height=10}
# plots to inspect the chosen model
plotVarianceExplained(proteomics_model_noDrug)
plotFactorCor(proteomics_model_noDrug)

for (data_set in names(proteomics_noDrug_obj@InputData)){
  plotWeightsHeatmap(
          proteomics_model_noDrug, 
          view = data_set, 
          factors = 1:proteomics_model_noDrug@Dimensions$K,
          show_colnames = FALSE)
    print(data_set)
}
```

```{r temporary, echo=FALSE, message=FALSE, warning=FALSE}
load(file.path(getwd(),"data/multiomics_MAE.RData"))

# change β to beta in rownames &
# set threshold of 1.5 and replace values above with NA
# apply rowMeans and rowVars cut offs

drug_resp <- assays(multiomics_MAE)$drug_resp_mono
rownames(drug_resp) <- gsub("\u03b2", "beta", rownames(drug_resp))
drug_resp[drug_resp >= 1.5] <- NA
drug_resp$rowvars <-  rowVars(as.matrix(drug_resp), na.rm = TRUE)
drug_resp$rowmeans <- rowMeans(as.matrix(drug_resp), na.rm = TRUE)

drug_resp <- drug_resp %>% rownames_to_column() %>% 
                   filter(rowmeans > 0.3) 
drug_resp <- drug_resp %>% 
                    filter(between(rowmeans, 0.9, 1.5)) %>%
                    filter(rowvars <= 0.02) %>% 
                    anti_join(drug_resp,.,)
drug_resp <- drug_resp %>% select(-c("rowvars", "rowmeans")) %>% 
                    column_to_rownames()
multiomics_MAE@ExperimentList$drug_resp_mono <- drug_resp
```


### IGHV & trisomy12
Factor loadings from patients on factors which are dominated by IGHV mutation status and trisomy12 are ploted against each other.
```{r factor scatter plot no drug}
plotFactorComparison(factors = c("LF1", "LF2"),
                     model_factors = proteomics_model_noDrug_factors)
```

### Top 10 weights
Top 10 features with the highest/lowest loadings are depicted per data view and factor.   

#### genetics
```{r}
lapply(c( 1:proteomics_model_noDrug@Dimensions$K), function(factor_num){
           plotTopWeights(proteomics_model_noDrug, 
                          view= "genetics", 
                          factor = factor_num)})
```

#### proteomics
```{r}
lapply(c( 1:proteomics_model_noDrug@Dimensions$K), function(factor_num){
           plotTopWeights(proteomics_model_noDrug, 
                          view= "proteomics", 
                          factor = factor_num)})
```

#### RNA seq
```{r}
lapply(c( 1:proteomics_model_noDrug@Dimensions$K), function(factor_num){
           plotTopWeights(proteomics_model_noDrug, 
                          view= "RNAseq_norm", 
                          factor = factor_num)})
```

### Enrichment analysis
```{r general parameter, model without drug resp}
views <- c("proteomics", "RNAseq_norm")

# term enrichment is performed for all factors in the model
proteomics_factors <- c(1:proteomics_model_noDrug@Dimensions$K)
RNAseq_factors <- c(1:proteomics_model_noDrug@Dimensions$K)
```

#### GO term enrichment
```{r }
FDR <- 0.1 # preliminary FDR filter
proteomics_model_gsa <- lapply(views, function(view){
  gsea <- runEnrichmentAnalysis(
    proteomics_model_noDrug,
    view = view,
    factors = "all",
    local.statistic = "loading",
    global.statistic = "mean.diff",
    statistical.test = "parametric",
    feature.sets = GO_sets,
    alpha = FDR
  )
  return(gsea)
})

# save enrichments
save(proteomics_model_gsa, file = paste0(model_output_path,
                                         "/proteomics_model_noDrug_GOgsea.RData"))


print(views[[1]])
plotEnrichmentBars(proteomics_model_gsa[[1]], alpha=FDR)

print(views[[2]])
plotEnrichmentBars(proteomics_model_gsa[[2]], alpha=FDR)

# select interesting factors with sign. pathways
proteomics_factors_keep <- lapply(proteomics_model_gsa[[1]]$sigPathways, 
                                  function(ls){
                                    ifelse(length(ls) != 0 , TRUE, FALSE)
                                  }) %>% unlist 
proteomics_factors_keep <- intersect(proteomics_factors,
                             c(1:length(proteomics_model_gsa[[1]]$sigPathways)
                               )[proteomics_factors_keep]) 


RNAseq_factors_keep <- lapply(proteomics_model_gsa[[2]]$sigPathways, 
                              function(ls){
                                ifelse(length(ls) != 0 , TRUE, FALSE)
                              }) %>% unlist
RNAseq_factors_keep <- intersect(RNAseq_factors,
                                 c(1:length(proteomics_model_gsa[[2]]$sigPathways)
                                 )[RNAseq_factors_keep]) 


```

##### Proteomics
```{r, fig.width=15, fig.height=13 }
lapply(proteomics_model_gsa[[1]]$sigPathways, as.data.frame)

fseaplots <- lapply(proteomics_factors_keep, function(factor) {
  plotEnrichmentDetailed(
    proteomics_model_noDrug,
    fsea.results =  proteomics_model_gsa[[1]],
    factor = factor,
    alpha = FDR,
    max.pathways = 15, # The top number of pathways to display
    feature.sets = GO_sets,
    max.genes = 10
  )
})

lapply(c(1:length(proteomics_factors_keep)), function(i){
  cowplot::plot_grid(fseaplots[[i]], labels = paste("proteomics factor",
                                                    proteomics_factors_keep[[i]]))
})
```

##### Enrichment barplot LF9
```{r, plot GO term enrichment LF9}
# plotting parameter
cex_text_plot <- 8/3

# plot enriched GO terms in top loading proteins of LF9
GO_LF9_bar_plot <-  proteomics_model_gsa[[1]]$pval.adj %>%
  as.data.frame() %>% rownames_to_column() %>%
  filter(rowname %in% proteomics_model_gsa[[1]]$sigPathways[[9]]) %>%
  dplyr::select(rowname, LF9) %>%   arrange(LF9) %>%
  mutate(pathway = as.factor(rowname)) %>%
  mutate(pathway = factor(pathway, levels = rev(pathway))) %>%
  ggplot(aes(pathway, -log10(LF9), fill=grepl("RNA", pathway ) )) +
  geom_col() + coord_flip() + pp_sra +
  theme(axis.title.y = element_blank()) +
  ylab("-log10(p-value enrichment)") +
  scale_fill_manual(values = c("grey", colors_CCP[5])) +
  annotate(geom="text", x=2.5, y=5, label="RNA Splicing",
           color=colors_CCP[5], size= cex_text_plot) +
  theme(legend.position = 'none')

GO_LF9_bar_plot
```


### Factor-Drug response correlation
```{r factor-response correlation, message=FALSE, warning=FALSE}
# retrieve loadings 
loadings <- proteomics_model_noDrug_factors %>% as.data.frame() %>%
  rownames_to_column(var = "primary")

# prepare drug response df
drug_resp <- longFormat(multiomics_MAE[,,"drug_resp_mono"]) %>%
                as.data.frame()
drug_resp <- drug_resp %>% filter( !( grepl("DMSO", rowname) | 
                                      grepl(" \\+ ", rowname)  |
                                      grepl("SSZ", rowname) | 
                                      grepl("phenyl",rowname) | 
                                      grepl("Vitamin D",rowname) ) ) 
drug_resp <- drug_resp[,c("primary", "rowname","value")] %>% 
              spread(rowname, value)

# order dfs based on patient IDs
drug_resp <- drug_resp[order(drug_resp$primary),]
loadings <- loadings[order(loadings$primary),]

# calculate correlations
drug_factor_cor <- lapply(colnames(loadings)[-1], function(factor_name){
    lapply(colnames(drug_resp[-1]), function(drug){
      ct <- cor.test(drug_resp[,drug], loadings[,factor_name],
                     method = "kendall")
      return(c("drug"=drug, "factor"=factor_name, "p_value"=ct$p.value,
             ct$estimate))
    }) %>% do.call(bind_rows,.)
  }) %>% do.call(bind_rows,.) %>% mutate(p_value=as.numeric(p_value), 
                                       tau=as.numeric(tau))

# p-value adjustments and annotation
drug_correlations <- drug_factor_cor %>% 
                      mutate(padj=p.adjust(p_value, method = "BH")) %>%
                      mutate(drug_name=gsub("\\_[1,2,3]", "", drug) ) %>% 
                      dplyr::filter(padj<0.1) %>%
                      left_join(.,
                        multiomics_MAE@metadata$drugs_functional_groups, 
                        by = c("drug_name"="Drug")) %>%
                      select(-drug_name)
                    


DT::datatable(drug_correlations)
```

# Session Info
```{r session info}
sessionInfo()
```
