---
title: "TTFT Trisomy12 and IGHV"
author: "Sophie Herbst & Mattias Vesterlund"
output:
  BiocStyle::html_document:
    self_contained: true
    code_download: true
    code_folding: hide
    toc: yes
    toc_depth: 4
    toc_float: yes
    df_print: paged
editor_options:
  chunk_output_type: console
---
Analysis date: `r Sys.Date()`


# Setup
## Load libraries
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(limma)
library(ggbeeswarm)
library(ggpubr)
library(readxl)
library(glmnet)
library(survival)
library(survminer)
require(switchBox)
require(gridExtra)
library(caret)
library(MLmetrics)
library(caret)
library(MLmetrics)
select <- dplyr::select

set.seed(2020)
```

## Load data
```{r, warning=FALSE, message=FALSE}
source("Data/Figure_layouts.R") 
DIA <- readRDS("Robjects/DIA2_alldata.RData")
load("Data/CLL_Proteomics_Setup.RData")
load("Data/multiomics_MAE.RData")
load("Data/CLL_Proteomics_ConsensusClustering.RData")
```

## Format data
```{r, warning=FALSE}
meta_Germany <- read_excel("Data/20200721_Patient_Annotation_health_records_pseud.xlsx",
  col_types = c( rep("text", 6), "skip", "text", rep("numeric", 7),  rep("skip", 4), "text", "date", "text",
                rep("date", 4), rep("text", 4), "date", "numeric", "date", "skip",
                "skip", "skip", "skip")
  )

patient_lookup_germany <- read_delim("Data/Patient_Lookup_IDs.txt", delim = "\t")

meta_Germany <- left_join(meta_Germany, patient_lookup_germany %>% select(-patient_ID_OMZ_first), by= c( "OMZ_Pat_ID", "NCT_Pat_ID"  ) )

meta_Germany$PG <- as.factor(CCP_group5[meta_Germany$patient_ID_CLL])

meta_HiRIEF <- meta_Germany %>%
  filter(!is.na(PG), diagnosis == "CLL") %>%
  select(patient_ID_CLL, PG, sampleDate, date_follow_up, alive, date_death, treatment_status )

meta_HiRIEF <- meta_HiRIEF %>%
  mutate(
    OS = difftime( date_death, sampleDate, units = "weeks" ),
    ObsTime_OS = if_else( alive == 0, OS ,
           difftime( date_follow_up, sampleDate, units = "weeks" ) ) )
```

### Prepare TTT data frame
```{r}
df_TTT <- DIA %>% 
  filter(grepl("Germany", cohort ) ) %>%
  select(Sample:ID, Pat_ID, date_next_treatment, date_death, alive, sampleDate, date_follow_up, cohort, treatment_at_collection) %>%
  unique %>%
  # Remove without any follow up or patients who were collected during treatment
  filter(is.na(treatment_at_collection) | treatment_at_collection!="intreatment") %>%
  mutate(TTT = difftime(date_next_treatment, sampleDate, units = "weeks"),
         ObsTime_TTT = if_else(alive==0, difftime(date_death,  sampleDate, units = "weeks"),
                                      difftime(date_follow_up, sampleDate, units = "weeks" )) ,
         received_next_treatment = if_else(is.na(TTT ), 0, 1 )) 



# Status: 1 for patients w/o next treatment ('censored'), 2 for patients who actually recieved treatment
df_TTT <- 
  df_TTT %>% 
  mutate(time = ifelse(received_next_treatment ==  0, ObsTime_TTT, 
                       if_else(received_next_treatment ==  1, TTT, as.difftime("NA")) ), 
         status = ifelse(received_next_treatment == 0, 1, 
                         if_else(received_next_treatment ==  1, 2, as.double(NA) ) ) )


```

### Prepare OS data frame from diagnosis
```{r}
df_OS <- DIA %>% 
  select(Sample:ID, Pat_ID, alive, sampleDate, cohort, OS, ObsTime_OS) %>%
  unique 


# Status: 1 for patients w/o next treatment ('censored'), 2 for patients who actually recieved treatment
df_OS <- 
  df_OS %>% 
  mutate(time = ifelse(alive ==  0, OS , 
                       if_else(alive ==  1, ObsTime_OS, as.difftime("NA")) ), 
         status = ifelse(alive == 1, 1, 
                         if_else(alive ==  0, 2, as.double(NA) ) ) )
```

### Prepare OS data frame from sample date
```{r}
df_OS_sampleDate <- DIA %>% 
  select(Sample:ID, Pat_ID, alive, sampleDate, cohort, OS = OS_since_sample, ObsTime_OS = ObsTime_OS_since_sample) %>%
  unique 


# Status: 1 for patients w/o next treatment ('censored'), 2 for patients who actually recieved treatment
df_OS_sampleDate <- 
  df_OS_sampleDate %>% 
  mutate(time = ifelse(alive ==  0, OS , 
                       if_else(alive ==  1, ObsTime_OS, as.difftime("NA")) ), 
         status = ifelse(alive == 1, 1, 
                         if_else(alive ==  0, 2, as.double(NA) ) ) )
```

### Prepare TTFT data frame
```{r}
df_TTFT <- DIA %>% 
  select(Sample:ID, Pat_ID, TTFT, ObsTime_TTFT, cohort, treatment_status, treatment_at_collection) %>%
  unique

# Status: 1 for patients w/o next treatment ('censored'), 2 for patients who actually recieved treatment
df_TTFT <- 
  df_TTFT %>% 
  mutate(time = ifelse(treatment_status ==  0, ObsTime_TTFT, 
                       if_else(treatment_status ==  1, TTFT, as.difftime("NA")) ), 
         status = ifelse(treatment_status == 0, 1, 
                         if_else(treatment_status ==  1, 2, as.double(NA) ) ) )

# Remove without any follow up or patients who were collected during treatment
df_TTFT <- filter(df_TTFT, ObsTime_TTFT!=0) %>% 
  filter(is.na(treatment_at_collection) | treatment_at_collection!="intreatment")
```


# Analysis
## Time to first treatment
```{r TTFT}
TTFT_tris_IGHV <- left_join(
  DIA %>% 
    select(Sample, IGHV_mutated, trisomy12, cohort) %>% 
    unique,
  df_TTFT,
  by=c("Sample", "cohort"))

df_TTFT_surv <- TTFT_tris_IGHV %>%
    filter(!cohort %in% c("Germany_1") ) %>%
    filter( !is.na(time), !is.na( status), !is.na(IGHV_mutated), !is.na(trisomy12) ) %>% 
    select(Sample,time, status, IGHV_mutated, trisomy12) %>%
    mutate( IGHV_str = if_else(IGHV_mutated == 1, "M-CLL",
                              if_else(IGHV_mutated == 0, "U-CLL", "NA") ),
            tris12_str = if_else(trisomy12 == 1, " & trisomy 12",
                                if_else(trisomy12 == 0, "", "NA"))) %>%
    mutate(IGHV_trisomy12 = paste0(IGHV_str, tris12_str) ) %>%
    unique() %>%
    dplyr::select(time, status, Factor = IGHV_trisomy12) %>%
    mutate(BinFactor = Factor)

df_TTFT_surv$Factor %>% table
  
formel <- as.formula(paste0("Surv(time, status) ~ Factor"))
  
# Years instead of months
df_TTFT_surv$time <- df_TTFT_surv$time/52.1428571
  
survfit_TTFT <- survfit(Surv(time, status) ~ BinFactor, data = df_TTFT_surv)
pp_sra_TTFT <- pp_sra
pp_sra_TTFT[[2]]$aspect.ratio <- 1
pp_sra_TTFT[[2]]$text <- element_text(size=15)
ggsurvplot(survfit_TTFT, data = df_TTFT_surv, 
                 xlab = "Time in years",
                 palette = c( colors_CCP[c(3, 1 ,4,2)]) ,
                 legend.labs = gsub("BinFactor=", "", names(survfit_TTFT$strata) ),
                 legend.title = "",
                 legend = "right",
                 ggtheme = pp_sra,
                 #xlim=c(0,7),
                 ylab="Treatment free survival",
                 pval = TRUE,
                 pval.method = TRUE,
                 conf.int = FALSE
                 ) + ggtitle("TTFT, IGHV and trisomy12")

tris12_IGHV_TTFT_survplot <- ggsurvplot(survfit_TTFT, data = df_TTFT_surv, 
                 xlab = "Time in years",
                 palette = c( colors_CCP[c(3, 1 ,4,2)]) ,
                 legend.labs = gsub("BinFactor=", "", names(survfit_TTFT$strata) ),
                 legend.title = "",
                 legend = "right",
                 ggtheme = pp_sra,
                 #xlim=c(0,7),
                 ylab="Treatment free survival",
                 pval = FALSE,
                 pval.method = FALSE,
                 conf.int = FALSE
                 )

```

### Only U-CLL
```{r TTFT_U}
TTFT_U_tris_IGHV <- left_join(
  DIA %>% 
    select(Sample, IGHV_mutated, trisomy12, cohort) %>% 
    unique,
  df_TTFT,
  by=c("Sample", "cohort"))

df_TTFT_U_surv <- TTFT_U_tris_IGHV %>%
    filter(!cohort %in% c("Germany_1"), IGHV_mutated==0 ) %>%
    filter( !is.na(time), !is.na( status), !is.na(IGHV_mutated), !is.na(trisomy12) ) %>% 
    select(Sample,time, status, IGHV_mutated, trisomy12) %>%
    mutate(IGHV.tris = interaction(IGHV_mutated, trisomy12)) %>%
    unique() %>%
    dplyr::select(time, status, Factor = IGHV.tris) %>%
    mutate(BinFactor = Factor)

df_TTFT_U_surv$Factor %>% table
  
formel <- as.formula(paste0("Surv(time, status) ~ Factor"))
  
# Years instead of months
df_TTFT_U_surv$time <- df_TTFT_U_surv$time/52.1428571
  
survfit_TTFT_U <- survfit(Surv(time, status) ~ BinFactor, data = df_TTFT_U_surv)
pp_sra_TTFT_U <- pp_sra
pp_sra_TTFT_U[[2]]$aspect.ratio <- 1
pp_sra_TTFT_U[[2]]$text <- element_text(size=15)
TTFT_U_survplot <- ggsurvplot(survfit_TTFT_U, data = df_TTFT_U_surv, 
                 xlab = "Time in years",
                 palette = c( colors_CCP[c(4,2)]) ,
                 legend.labs = gsub("BinFactor=", "", names(survfit_TTFT_U$strata) ),
                 legend.title = "IGHV.trisomy12",
                 legend = "right",
                 ggtheme = pp_sra,
                 #xlim=c(0,7),
                 ylab="Treatment free survival",
                 pval = TRUE,
                 pval.method = TRUE,
                 conf.int = FALSE
                 ) 

TTFT_U_survplot + ggtitle("TTFT U-CLL, IGHV and trisomy12")

```

### Only M-CLL
```{r TTFT_M}
TTFT_M_tris_IGHV <- left_join(
  DIA %>% 
    select(Sample, IGHV_mutated, trisomy12, cohort) %>% 
    unique,
  df_TTFT,
  by=c("Sample", "cohort"))

df_TTFT_M_surv <- TTFT_M_tris_IGHV %>%
    filter(!cohort %in% c("Germany_1"), IGHV_mutated==1 ) %>%
    filter( !is.na(time), !is.na( status), !is.na(IGHV_mutated), !is.na(trisomy12) ) %>% 
    select(Sample,time, status, IGHV_mutated, trisomy12) %>%
    mutate(IGHV.tris = interaction(IGHV_mutated, trisomy12)) %>%
    unique() %>%
    dplyr::select(time, status, Factor = IGHV.tris) %>%
    mutate(BinFactor = Factor)

df_TTFT_M_surv$Factor %>% table
  
formel <- as.formula(paste0("Surv(time, status) ~ Factor"))
  
# Years instead of months
df_TTFT_M_surv$time <- df_TTFT_M_surv$time/52.1428571
  
survfit_TTFT_M <- survfit(Surv(time, status) ~ BinFactor, data = df_TTFT_M_surv)
pp_sra_TTFT_M <- pp_sra
pp_sra_TTFT_M[[2]]$aspect.ratio <- 1
pp_sra_TTFT_M[[2]]$text <- element_text(size=15)
TTFT_M_survplot <- ggsurvplot(survfit_TTFT_M, data = df_TTFT_M_surv, 
                 xlab = "Time in years",
                 palette = c( colors_CCP[c( 3 ,1)]) ,
                 legend.labs = gsub("BinFactor=", "", names(survfit_TTFT_M$strata) ),
                 legend.title = "IGHV.trisomy12",
                 legend = "right",
                 ggtheme = pp_sra,
                 #xlim=c(0,7),
                 ylab="Treatment free survival",
                 pval = TRUE,
                 pval.method = TRUE,
                 conf.int = FALSE
                 ) 

TTFT_M_survplot + ggtitle("TTFT M-CLL, IGHV and trisomy12")

```

# Conclusion
M-CLL trisomy 12 patients have shorter TTFT than M-CLL patients without trisomy 12

# Save important objects
```{r}
```

# Save important plots
```{r}
save(tris12_IGHV_TTFT_survplot,
     file = "Robjects/TTFT_IGHV_tris12_plot.RData")
```

# Session Info
```{r}
sessionInfo()
```

```{r knitr exit}
knitr::knit_exit()
```

